Resolvable Ambiguity

### 2019-07-11 paper/paper.tex
Pass over beginning of 6.3

This section introduces a linear encoding of the lattices of the previous chapter as words. Lemmas~\ref{lemma:same-basic} and \ref{lemma:top-bottom-determine} imply that a lattice encoding requires three things: a basic word, a set of \emph{required} parentheses, and a set of \emph{possible} parentheses. We encode required parentheses with ``$\reqp{}$'' and possible parentheses with ``$\posp{}$''. For example, the lattice in Fig.~\ref{fig:lattice2} is represented by ``$\posp{\posp{1} + \posp{\posp{2} + \posp{3}}}$'', while the tree for ``$(1 + 2) * 3$'' has a lattice represented by ``$\posp{\reqp{\posp{1} + \posp{2}} * \posp{3}}$''. With this encoding we can apply the formidable body of knowledge available for word languages, albeit with some extra care; the linear encoding of a lattice is not unique.

For example, ``$\posp{\posp{1}}$'' and ``$\posp{1}$'' represent the same lattice, as do ``$\reqp{\posp{2}}$'' and ``$\reqp{2}$''. To gain uniqueness we forbid duplicated parentheses and prioritize required parentheses over possible parentheses, e.g., ``$\posp{\posp{1}}$'' is uniquely represented as ``$\posp{1}$'' while ``$\reqp{\posp{2}}$'' is uniquely represented as ``$\reqp{2}$''.

Lattice equality is now equvialent with equality of the linear encoding. To determine if a lattice is entirely contained in another, we make the following observation: we can move the top of the lattice ``up'' (new top rangeset is a strict superset) by adding a pair of possible parentheses, and move the bottom ``down'' (new bottom rangeset is a strict subset) by replacing a required pair with a possible pair. For example, ``$\reqp{1}2$'' is entirely contained in ``$\reqp{1}\posp{2}$'', which is entirely contained in ``$\posp{1}\posp{2}$''.

The centerpiece of our algorithm is a visibly pushdown automaton constructed in such a way that:

\begin{itemize}
\item There is a bijection between trees in $L(T_D)$ and successful runs.
\item The word recognized by a successful run is the linear encoding of the corresponding tree's lattice.
\end{itemize}

>>>>

This section introduces a linear encoding of the lattices of the previous section as words. Lemmas~\ref{lemma:same-basic} and \ref{lemma:top-bottom-determine} imply that a lattice encoding requires three things: a basic word, a set of \emph{required} parentheses, and a set of \emph{possible} parentheses. We encode required parentheses with ``$\reqp{}$'' and possible parentheses with ``$\posp{}$''. For example, the lattice in Fig.~\ref{fig:lattice2} is represented by ``$\posp{\posp{1} + \posp{\posp{2} + \posp{3}}}$'', while the tree for ``$(1 + 2) * 3$'' has a lattice represented by ``$\posp{\reqp{\posp{1} + \posp{2}} * \posp{3}}$''. With this encoding we can apply the formidable body of knowledge available for word languages, albeit with some extra care; the linear encoding of a lattice is not unique.

For example, ``$\posp{\posp{1}}$'' and ``$\posp{1}$'' represent the same lattice, as do ``$\reqp{\posp{2}}$'' and ``$\reqp{2}$''. To gain uniqueness we forbid duplicated parentheses and prioritize required parentheses over possible parentheses. For example, ``$\posp{\posp{1}}$'' is uniquely represented as ``$\posp{1}$'' while ``$\reqp{\posp{2}}$'' is uniquely represented as ``$\reqp{2}$''.

Lattice equality is now equivalent with equality of the linear encoding. To determine if a lattice is entirely contained in another, we make the following observation: we can move the top of the lattice ``up'' (new top rangeset is a strict superset) by adding a pair of possible parentheses, and move the bottom ``down'' (new bottom rangeset is a strict subset) by replacing a required pair with a possible pair. For example, ``$\reqp{1}2$'' is entirely contained in ``$\reqp{1}\posp{2}$'', which is entirely contained in ``$\posp{1}\posp{2}$''.

The centerpiece of our algorithm is a \emph{visibly pushdown automaton} constructed in such a way that

\begin{itemize}
\item there is a bijection between trees in $L(T_D)$ and successful runs; and
\item the word recognized by a successful run is the linear encoding of the corresponding tree's lattice.
\end{itemize}
